---
title: "p8105_hw2_ss5929"
author: "Shuyi SHEN"
date: "9/24/2019"
output: github_document
---
## Problem 1
```{r}
library(tidyverse)
library(readxl)

TrashData = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                       sheet = 1,
                       range = "A2:N408"
                       ) %>%
           janitor::clean_names() %>%
           drop_na  (dumpster) %>%
           mutate_at(vars(sports_balls),round,0) 
         
pre_2017 = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                       sheet = 6,
                       skip  = 1) %>%  
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate( year= "2017" ) 

pre_2018 = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
                       sheet = 5,
                       skip  = 1) %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate( year= "2018" ) 

#combine two precipitation datasets
pre_da = rbind(pre_2017,pre_2018) %>%
mutate(month = month.name[month]) 

  
```

**Summary**

*TrashData*

There are `r nrow(TrashData)` observations in the `TrashData`dataset  with `r ncol(TrashData)` variables including information on the dumpster number,date of collection(``r colnames(TrashData)[2:4]``),amount of total litter(``r colnames(TrashData)[5:6]``), litter type(``r colnames(TrashData)[7:13]``) and ``r colnames(TrashData)[14]``(each ton of trash equates to on average 500 kilowatts of electricity and an average household will use 30 kilowatts per day ). 

The key variables are ``r colnames(TrashData)[5:14]``.

*pre_2017*

There are `r nrow(pre_2017)` observations in the `pre_2017` dataset with `r ncol(pre_2017)` variables including ``r colnames(pre_2017)``. The key variable is ``r colnames(pre_2017)[2]``, which means the total precipitation in that month.



*pre_2018*

There are `r nrow(pre_2018)` observations in the `pre_2018` dataset with `r ncol(pre_2018)` variables including ``r colnames(pre_2018)``. The key variable is ``r colnames(pre_2018)[2]``, which means the total precipitation in that month.


*pre_da*

There are `r nrow(pre_da)` observations with `r ncol(pre_da)` variables including ``r colnames(pre_da)`` in the `pre_da` dataset. The key variable is ``r colnames(pre_da)[2]``, which means the total precipitation in that month that year.


The total precipitation in 2018 is `r sum(pre_2018$total)`.

The median number of sports balls in a dumpster in 2017 is `r median(TrashData$sports_balls) `.

## Problem 2
```{r message= FALSE}
pols<-read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>%
   janitor::clean_names() %>%
   separate(mon,sep="-",into = c("year","month","day")) %>%  
   mutate(month=as.numeric(month))%>%
   mutate(month=month.name[month])%>%
   pivot_longer(
    cols         = starts_with("prez"),
    names_to     = "president",
    names_prefix = "prez_"
   )%>% 
   filter(value!="0") %>%
   select(-value) %>%
   select(-day) 

snp<-read_csv("data/fivethirtyeight_datasets/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date,sep="/",into = c("month","day","year")) %>% 
   mutate (month=as.numeric(month))%>%
   arrange(year,month) %>% 
   mutate (month=month.name[month]) %>%
   select (year,month, close) 

# merge the first and second data set
pols_snp <- left_join(pols,snp,by=c("year","month"))

unemployment<-read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>%
  janitor::clean_names()%>%
  pivot_longer(
     jan:dec,
     names_to = "month" 
) %>% 
mutate(year  = as.character(year)) %>%
mutate(month = factor(month)) %>%
mutate(month = month.name[month])   

#merge all three data set
result<-left_join(pols_snp,unemployment,by=c("year","month"))


```

**Summary**

*pols*

There are `r nrow(pols)` objects with `r ncol(pols)` variables in `pols` dataset. The variables are:

``r colnames(pols)[1]``: the associated year

``r colnames(pols)[2]``: the associated month

``r colnames(pols)[3]``: the number of republican governors on the associated date

``r colnames(pols)[4]``: the number of republican senators on the associated date

``r colnames(pols)[5]``: the number of republican representatives on the associated date

``r colnames(pols)[6]``: the number of democratic governors on the associated date

``r colnames(pols)[7]``: the number of democratic senators on the associated date

``r colnames(pols)[8]``: the number of democratic representatives on the associated date

``r colnames(pols)[9]``: whether the president was republican or democratic on the associated date

and key variables are ``r colnames(pols)[3:9]`` 

The range of years is from `r range(pols$year)[1]` to `r range(pols$year)[2]`. 

*snp*

There are `r nrow(snp)` objects with `r ncol(snp)` variables in `snp` dataset. The variables are ``r colnames(snp)`` and key variables is ``r colnames(snp)[3]`` indicating the closing values of the S&P stock index on the associated date

The range of years is from `r range(snp$year)[1]` to 
`r range(snp$year)[2]`.

*unemployment*

There are `r nrow(unemployment)` objects with `r ncol(unemployment)` variables in `umemployment` dataset. The variables are ``r colnames(unemployment)`` and key variables is ``r colnames(unemployment)[3]`` indicating percentage of unemployment in that month in that year. 

The range of years is from `r range(unemployment$year)[1]` to `r range(unemployment$year)[2]`.

*pols_snp*

There are `r nrow(pols_snp)` objects with `r ncol(pols_snp)` variables in `pols_snp` dataset. The variables are ``r colnames(pols_snp)`` and key variables are ``r colnames(pols_snp)[3:10]``.

*result*

There are `r nrow(result)` objects with `r ncol(result)` variables in `result` dataset. The variables are ``r colnames(result)``.

The range of years is from `r range(result$year)[1]` to `r range(result$year)[2]` .

## Problem 3
```{r message=FALSE}
library(ggplot2)

# read the data,clean and tidy
BabyNames = read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  distinct()%>% 
  mutate(childs_first_name=str_to_lower(childs_first_name)) %>%
  mutate(
    ethnicity=recode(ethnicity,"WHITE NON HISP"="WHITE NON HISPANIC",
                               "BLACK NON HISP"="BLACK NON HISPANIC",
                               "ASIAN AND PACI"="ASIAN AND PACIFIC ISLANDER")
  ) 

# the popularity of olivia among different ethnicity over time  
Olivia=BabyNames %>%
filter(childs_first_name=="olivia")   
 tibble(
    year      = Olivia$year_of_birth,
    ethnicity = Olivia$ethnicity,
    rank      = Olivia$rank
  ) %>% 
  pivot_wider(
    names_from  = "year",
    values_from = "rank" 
  ) %>% 
      knitr::kable(format="html")

# the most popular name among the male in different ethnicity over time  
names_male=BabyNames %>% 
  filter(gender == "MALE",rank==1) %>% 
  select(year_of_birth,childs_first_name,ethnicity) %>% 
  pivot_wider(
    names_from  = "year_of_birth",
    values_from = "childs_first_name"
  ) 
names_male %>% knitr::kable(format = "html")


# plot the relationship between rank and count among male in white non-hispanic
BabyNames %>%
  filter(gender=="MALE",
         ethnicity=="WHITE NON HISPANIC", 
         year_of_birth=="2016") %>% 
  ggplot     (aes(x=rank,y=count))+
  geom_point (aes(size=rank),alpha=.5,color="blue")+
  labs       (title = "Count versus Rank", subtitle="Among Male in White Non Hispanic in 2016")+
  theme      (plot.title=element_text(hjust=0.5),plot.subtitle = element_text(hjust = 0.5))

```
**Summary**

*BabyNames*

There are `r nrow(BabyNames)` observations with `r ncol(BabyNames)` variables in the dataset. The variables are 
``r colnames(BabyNames)[1]``: year of birth

``r colnames(BabyNames)[2]``: sex of the baby

``r colnames(BabyNames)[3]``: mother's ethnicity

``r colnames(BabyNames)[4]``: child's first name

``r colnames(BabyNames)[5]``: number of babies with this name

``r colnames(BabyNames)[6]``: frequency of baby names in descending order

and key variables are ``r colnames(BabyNames)[4:6]``

*Olivia*

There are 3 varibles in this dataset including `year`,`ethnicity` and `rank` indicating the rank of Olivia among female in different ethnicity in different years.

*names_male*

There are 3 variables in this dataset including `year`,`ethnicity` and `childs_first_name` indicating the most popular name among male in different ethnicity over time.

*plot*

The graph shows the increase of count with the decrease of rank among male in white non hispanic in 2016. The xlab is rank while the ylab is count and the point means one specific childs_first_name. The size of point indicates the rank.


 


